---
title: "Gateway_library"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(RColorBrewer)
```

## Reading in the csv
Read in the csv with all of the (-), (+), and fold induction values
get rid of all extra columns (there were none in this csv)
remove gfpUV (won't want this control for the graph)
put NA in all of the places where there isn't data 
remove rows that only have NA in them 
create separate charts for each variable 

```{r}
gateway_library <- read.csv("gateway_library_data.csv", header=T)
gateway_library <- subset.data.frame(gateway_library, select = -gfpUV)
#gateway_library[,2:ncol(gateway_library)] <- as.numeric(gateway_library[,2:ncol], na.ram=TRUE)
gateway_library[gateway_library == -2] <-NA
gateway_repression1 <- gateway_library[gateway_library == "(-)",]
gateway_repression <- gateway_repression1[rowSums(is.na(gateway_repression1)) !=ncol(gateway_repression1),]
rm(gateway_repression1)
gateway_expression1 <- gateway_library[gateway_library == "(+)",]
gateway_expression <- gateway_expression1[rowSums(is.na(gateway_expression1)) !=ncol(gateway_expression1),]
rm(gateway_expression1)
gateway_fold_induction1 <- gateway_library[gateway_library == "FI",]
gateway_fold_induction <- gateway_fold_induction1[rowSums(is.na(gateway_fold_induction1)) !=ncol(gateway_fold_induction1),]
rm(gateway_fold_induction1)
```

## Statistics Time Baby


```{r}
summary(gateway_repression)
summary(gateway_expression)
summary(gateway_fold_induction)
```

##long form baby
take each dataframe and turn it into long form from wide form
the second line for each is to take the fluoresence values (which R took as characters) and make them numerical
```{r}
long_gateway_library <- gather(gateway_library,Colony_Name, Fluorescence, -COLONY)
long_gateway_library[,3] <- as.numeric(long_gateway_library[,3], na.rm=TRUE)
long_gateway_repression <- gather(gateway_repression,Colony_Name, Fluorescence, -COLONY)
long_gateway_repression[,3] <- as.numeric(long_gateway_repression[,3], na.rm=TRUE)
long_gateway_expression <- gather(gateway_expression,Colony_Name, Fluorescence, -COLONY)
long_gateway_expression[,3] <- as.numeric(long_gateway_expression[,3], na.rm=TRUE)
long_gateway_fold_induction <- gather(gateway_fold_induction,Colony_Name, Fluorescence, -COLONY)
long_gateway_fold_induction[,3] <- as.numeric(long_gateway_fold_induction[,3], na.rm=TRUE)
long_gateway_rep_and_FI <-rbind(long_gateway_repression, long_gateway_fold_induction)
long_gateway_rep_and_exp <- rbind(long_gateway_repression, long_gateway_expression)
```


##Let's make some figures
```{r}
ggplot(long_gateway_rep_and_FI, aes(x=reorder(Colony_Name,-Fluorescence, na.rm=TRUE), y=Fluorescence, fill=COLONY))+
    geom_boxplot()+
   labs(title="Gateway Library Repression and Fold Induction", x="Colony", y="")+
   scale_y_log10()+
+     theme_classic()
ggplot(long_gateway_rep_and_exp, aes(x=reorder(Colony_Name,-Fluorescence, na.rm=TRUE), y=Fluorescence, fill=COLONY))+
     geom_boxplot()+
   labs(title="Gateway Library Repression and Expression", x="Colony", y="")+
    scale_y_log10()+
     theme_classic()
```

