---
title: "Loop library analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(RColorBrewer)
```

## Reading in the csv
Read in the csv with all of the (-), (+), and fold induction values
get rid of all extra columns 
remove gfpUV (won't want this control for the graph)
put NA in all of the places where there isn't data 
create separate charts for each variable 
```{r}
loop_library <- read.csv("loop_library_data.csv", header=T)
loop_library <- subset.data.frame(loop_library, select = -X)
loop_library <- subset.data.frame(loop_library, select = -gfpUV)
loop_library[loop_library == -2] <-NA
loop_repression <- loop_library[loop_library == "(-)",]
loop_expression <- loop_library[loop_library == "(+)",]
loop_fold_induction <- loop_library[loop_library == "FI",]
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
## Statistics Time Baby
```{r}
summary(loop_repression)
summary(loop_expression)
summary(loop_fold_induction)
```

##long form baby
```{r}
long_loop_library <- gather(loop_library,Colony_Name, Fluorescence, -COLONIES)
long_loop_repression <- gather(loop_repression,Colony_Name, Fluorescence, -COLONIES)
long_loop_expression <- gather(loop_expression,Colony_Name, Fluorescence, -COLONIES)
long_loop_fold_induction <- gather(loop_fold_induction,Colony_Name, Fluorescence, -COLONIES)
long_loop_rep_and_FI <-rbind(long_loop_repression, long_loop_fold_induction)
```


##Let's make some figures
```{r}
ggplot(long_loop_rep_and_FI, aes(x=reorder(Colony_Name,-Fluorescence, na.rm=TRUE), y=Fluorescence, fill=COLONIES))+
  geom_boxplot()+
  labs(title="Loop Library Repression and Fold Induction", x="Colony", y="")+
  scale_y_log10()+
  theme_classic()
```

