---
title: "stem8_stats"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(RColorBrewer)
library(plyr)
library(dplyr)
library(tibble)
```

## Creating the function to remove outliers based on Interquartile range
```{r}
removeOut <- function(distribution, bounds=1.5, na.rm = TRUE) {
 Q <- quantile(distribution, probs=c(.25,.75), na.rm=TRUE) 
 #finds the 25th and 75th percentile of the distribution, excluding NAs
 iqr <- IQR(distribution, na.rm=TRUE)
 #difference between 75th and 25th quartile=interquartile range
 upperBound <- Q[2]+bounds*iqr
 #upper bound of the values is 75th quartile + 1.5*(interquartile range) note: that 1.5 can be vaired by changing bound
 lowerBound <- Q[1]-bounds*iqr
 #upper bound of the values is 25th quartile - 1.5*(interquartile range)
 eliminated <- c(distribution < lowerBound | distribution > upperBound)
 #take out the values that are outside of the established bounds note: that 1.5 can be vaired by changing bound
 noteliminatedindex <- which(distribution > lowerBound & distribution < upperBound)
 #index values within the upper and lower bounds
 noteliminatedvalues <- distribution[noteliminatedindex]
 return(noteliminatedvalues)
 #return those values which will be included that are between the upper and lower bounds 
}
```

##Removing Outliers for Repression (measurements taken in the absence of ligand)
Read in csv.
Formatting of csv: each column is a different colony, first row is the name of the colony, each column is then filled with the background fluorescence corrected flourescence values (there should be three values per biological replicate, each of which are the technical replicates). Far right column is the condition of the growth, either - or + ligand. For this particular switch, -L=repression and +L=expression. For those colonies that have less raw data values than the controls, the empty cells are filled with an integer that can be removed easily (-2) to be turned into NA when inputted into R. 
Put NA in place of -2 (which was how NA was denoted in the csv).
This is for repression and expression, fold induction will not be included because it will be calculated based on the included repression and expression values.
```{r}
stem8_library <- read.csv("stem8_library_data.csv", header=T)
stem8_library[stem8_library == -2] <-NA
stem8_rep1 <- stem8_library[stem8_library == "(-)",]
#dataframe of just repression
stem8_rep <- stem8_rep1[rowSums(is.na(stem8_rep1)) !=ncol(stem8_rep1),]
rm(stem8_rep1)
#remove rows that were only NA and delete dataframe that was filled with NA only columns
#stem8_rep[,40] <- as.numeric(stem8_rep[,40], na.rm=TRUE)
#one of the columns was being read in as a character so this fixes that one

```

###Using function to remove outliers and create a new dataframe in wide form and long form for repression
```{r}

noOut_stem8_rep.list <- apply(stem8_rep[,2:ncol(stem8_rep)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_stem8_rep_sideways <- plyr::ldply(noOut_stem8_rep.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_stem8_rep <- data.frame(t(noOut_stem8_rep_sideways[-1]))
colnames(noOut_stem8_rep) <- noOut_stem8_rep_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
stem8_rep_noOut_wideform <- cbind(stem8_rep[,1], noOut_stem8_rep[,1:ncol(noOut_stem8_rep)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(stem8_rep_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
stem8_rep_noOut_wideform_noUV <- subset.data.frame(stem8_rep_noOut_wideform, select = -gfpUV)
#remove gfpUV --> this is because gfpUV should have outliers removed but it does not need to be considered in the analysis between and among groups since it is a negative control, keeping it in the dataset will give false significance 
long_stem8_rep <- gather(stem8_rep_noOut_wideform_noUV, Colony_Name, Fluorescence, -COLONIES)
is_NH5_stem8 <- vector()
# 0=not NH5 4=evolved NH5 8=control NH5
is_NH5_stem8_evolved <- grep("\\.NH5", long_stem8_rep[,2])
jj=1
for(ii in 1:nrow(long_stem8_rep)) {
  if(long_stem8_rep[ii,2] == "NH5") {
    is_NH5_stem8[ii] <- 8
  }
  else if(ii == is_NH5_stem8_evolved[jj]) {
    is_NH5_stem8[ii] <- 4
    jj = jj+1
    if(jj>27){
      jj = 1
    }
  }
  else{
    is_NH5_stem8[ii] <- 0
  }
}

long_stem8_rep <- cbind(long_stem8_rep, is_NH5_stem8)
#make column that identifies colonies as NH5 for plotting
```
 
##Removing Outliers for Expression (measurements taken in the presence of ligand)
```{r}
stem8_exp1 <- stem8_library[stem8_library == "(+)",]
#dataframe of just expression
stem8_exp <- stem8_exp1[rowSums(is.na(stem8_exp1)) !=ncol(stem8_exp1),]
rm(stem8_exp1)
#remove rows that were only NA 
stem8_exp[,40] <- as.numeric(stem8_exp[,40], na.rm=TRUE)
#one of the columns was being read in as a character so this fixes that one

```

###Using function to remove outliers and create a new dataframe in wide form and long form for expression
```{r}

noOut_stem8_exp.list <- apply(stem8_exp[,2:ncol(stem8_exp)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_stem8_exp_sideways <- plyr::ldply(noOut_stem8_exp.list, rbind)
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_stem8_exp <- data.frame(t(noOut_stem8_exp_sideways[-1]))
colnames(noOut_stem8_exp) <- noOut_stem8_exp_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
stem8_exp_noOut_wideform <- cbind(stem8_exp[1:78,1], noOut_stem8_exp[,1:ncol(noOut_stem8_exp)])
#replaces id column with COLONY column which identified the value as (-) or (+), did 1:78 because the noOut dataframe is shorter than the dataframe with outliers and the two need to have the same number of rows 
colnames(stem8_exp_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
stem8_exp_noOut_wideform_noUV <- subset.data.frame(stem8_exp_noOut_wideform, select = -gfpUV)
#remove gfpUV --> this is because gfpUV should have outliers removed but it does not need to be considered in the analysis between and among groups since it is a negative control, keeping it in the dataset will give false significance 
long_stem8_exp <- gather(stem8_exp_noOut_wideform_noUV, Colony_Name, Fluorescence, -COLONIES)
is_NH5_stem8exp <- vector()
for(ii in 1:nrow(long_stem8_exp)) {
  if(long_stem8_exp[ii,2] == "NH5") {
    is_NH5_stem8exp[ii] <- 1
  }
  else{
    is_NH5_stem8exp[ii] <- 0
  }
}
long_stem8_exp <- cbind(long_stem8_exp, is_NH5_stem8exp)
#make column that identifies colonies as NH5 for plotting
```


##ANOVA
###repression 
```{r, echo=FALSE}
stem8_rep.aov <- aov(Fluorescence ~ Colony_Name, data=long_stem8_rep)
summary(stem8_rep.aov)
colony_comparisons <- TukeyHSD(stem8_rep.aov)
Stem8_rep_confidence_interval <- as.data.frame(colony_comparisons[["Colony_Name"]])
Stem8_rep_confidence_interval <- rownames_to_column(Stem8_rep_confidence_interval)
Stem8_repvsNH5_confidence_interval <- Stem8_rep_confidence_interval[grep("NH5", Stem8_rep_confidence_interval[,1]),]
medianstem8rep <- as.data.frame(apply(noOut_stem8_rep, 2, median, na.rm=TRUE))
is_sigstem <- vector()
for(ii in 1:nrow(Stem8_repvsNH5_confidence_interval)) {
  if(Stem8_repvsNH5_confidence_interval[ii,5] < 0.001) {
    is_sigstem[ii] <- 3
  }
  else if(Stem8_repvsNH5_confidence_interval[ii,5] < 0.01) {
    is_sigstem[ii] <- 2
  }
  else if(Stem8_repvsNH5_confidence_interval[ii,5] < 0.05) {
    is_sigstem[ii] <- 1
  }
  else{
    is_sigstem[ii] <- 0
  }
}
Stem8_repvsNH5_confidence_interval <- cbind(Stem8_repvsNH5_confidence_interval, is_sigstem)

``` 
###expression
```{r, echo=FALSE}
stem8_exp.aov <- aov(Fluorescence ~ Colony_Name, data=long_stem8_exp)
summary(stem8_exp.aov)
colony_comparisons <- TukeyHSD(stem8_exp.aov)
Stem8_exp_confidence_interval <- as.data.frame(colony_comparisons[["Colony_Name"]])
Stem8_exp_confidence_interval <- rownames_to_column(Stem8_exp_confidence_interval)
Stem8_expvsNH5_confidence_interval <- Stem8_exp_confidence_interval[grep("NH5", Stem8_exp_confidence_interval[,1]),]
medianstem8exp <- as.data.frame(apply(noOut_stem8_exp, 2, median, na.rm=TRUE))

``` 
###fold induction

```{r}
Fold_induction_stem8 <- cbind(medianstem8exp,medianstem8rep)
colnames(Fold_induction_stem8) <- c("Expression", "Repression")
FI_stem8 <- Fold_induction_stem8
FI_stem8$FI_stem8 <- Fold_induction_stem8$Expression / Fold_induction_stem8$Repression
FI_stem8 <- tibble::rownames_to_column(FI_stem8, "Colony_Name")
``` 

##Plots
```{r, echo=FALSE}
ggplot(long_stem8_exp, aes(x=reorder(Colony_Name,-Fluorescence, na.rm=TRUE), y=Fluorescence, fill=is_NH5_stem8exp))+
    geom_boxplot(alpha=0.8)+
   labs(title="Hairpin Stem Library Expression", x="Colony", y="Fluorescence")+
   scale_y_log10()+
    theme_classic()+
    theme(axis.text.x = element_text(face="bold", angle=90), plot.title = element_text(hjust=0.5, face="bold"), axis.text.y = element_text(face="bold"), axis.title.x = element_text(face = "bold"), legend.position = "none")

ggplot(long_stem8_rep, aes(x=reorder(Colony_Name,-Fluorescence, na.rm=TRUE), y=Fluorescence, fill=is_NH5_stem8))+
    geom_boxplot(alpha=0.8)+
   labs(title="Hairpin Stem Library Repression", x="Colony", y="Fluorescence")+
   scale_y_log10()+
    theme_classic()+
    theme(plot.title = element_text(hjust=0.5, face="bold"), axis.text.y = element_text(face="bold"), axis.title.x = element_text(face = "bold"), axis.title.y = element_text(face = "bold"), legend.position = "none", axis.text.x=element_blank())

ggplot(FI_stem8, aes(x=reorder(Colony_Name,-FI_stem8), y=FI_stem8))+
    geom_point()+
   labs(title="Hairpin Stem Library Fold Induction", x="Colony", y="Fold Induction")+
 #  scale_y_log10()+
    theme_classic()+
    theme(axis.text.x = element_text(face="bold", angle=90), plot.title = element_text(hjust=0.5, face="bold"), axis.text.y = element_text(face="bold"), axis.title.x = element_text(face = "bold"), axis.title.y = element_text(face = "bold"), legend.position = "none")

``` 
