---
title: "anchor_exchange_analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(RColorBrewer)
library(plyr)
library(dplyr)
library(tibble)
library(ggsignif)
library(plotrix)
library(openxlsx)
library(viridis)
```

## Creating the function to remove outliers based on Interquartile range
```{r}
removeOut <- function(distribution, bounds=1.5, na.rm = TRUE) {
 Q <- quantile(distribution, probs=c(.25,.75), na.rm=TRUE) 
 #finds the 25th and 75th percentile of the distribution, excluding NAs
 iqr <- IQR(distribution, na.rm=TRUE)
 #difference between 75th and 25th quartile=interquartile range
 upperBound <- Q[2]+bounds*iqr
 #upper bound of the values is 75th quartile + 1.5*(interquartile range) note: that 1.5 can be varied by changing bound
 lowerBound <- Q[1]-bounds*iqr
 #upper bound of the values is 25th quartile - 1.5*(interquartile range)
 eliminated <- c(distribution < lowerBound | distribution > upperBound)
 #take out the values that are outside of the established bounds note: that 1.5 can be varied by changing bound
 noteliminatedindex <- which(distribution > lowerBound & distribution < upperBound)
 #index values within the upper and lower bounds
 noteliminatedvalues <- distribution[noteliminatedindex]
 return(noteliminatedvalues)
 #return those values which will be included that are between the upper and lower bounds 
}
```

Read in csv.
Formatting of csv: each column is a different colony, first row is the name of the colony, each column is then filled with the background fluorescence (PBR322) corrected fluorescence values (there should be three values per biological replicate, each of which are the technical replicates). Far right column is the condition of the growth, either - or + ligand. For this particular switch, -L=repression and +L=expression. For those colonies that have less raw data values than the controls, the empty cells are filled with an integer that can be removed easily (-2) to be turned into NA when read into R. 
This is for repression and expression, fold induction will not be included because it will be calculated based on the included repression and expression values.
```{r}
#read in csv, these values will include colonies below the function cut off which has been defined here as 2.0
  NH5_AEall <- read.csv("~/Downloads/Expression_platform_project/expression_platform_files/AE_DATA.csv", header=T)
empty_columns <- sapply(NH5_AEall, function(x) all(is.na(x) | x==""))
NH5_AEall <- NH5_AEall[,!empty_columns]
NH5_AEall[NH5_AEall == -2] <-NA
#MINUS 2AP DF
NH5_AE_minus1 <- NH5_AEall[NH5_AEall == "(-)",]
NH5_AE_minus <- NH5_AE_minus1[rowSums(is.na(NH5_AE_minus1)) !=ncol(NH5_AE_minus1),]
rm(NH5_AE_minus1)
#remove rows that were only NA and delete dataframe that was filled with NA only columns
#PLUS 2AP DF
NH5_AE_plus1 <- NH5_AEall[NH5_AEall == "(+)",]
NH5_AE_plus <- NH5_AE_plus1[rowSums(is.na(NH5_AE_plus1)) !=ncol(NH5_AE_plus1),]
rm(NH5_AE_plus1)
#remove rows that were only NA 
```

###Using function to remove outliers and create a new dataframe in wide form and long form for minus
```{r}
noOut_NH5_AE_minus.list <- apply(NH5_AE_minus[,2:ncol(NH5_AE_minus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_NH5_AE_minus_sideways <- plyr::ldply(noOut_NH5_AE_minus.list, rbind)
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_NH5_AE_minus <- data.frame(t(noOut_NH5_AE_minus_sideways[-1]))
colnames(noOut_NH5_AE_minus) <- noOut_NH5_AE_minus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
NH5_AE_minus_noOut_wideform <- cbind(NH5_AE_minus[1:nrow(noOut_NH5_AE_minus),1], noOut_NH5_AE_minus[,1:ncol(noOut_NH5_AE_minus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(NH5_AE_minus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_NH5_AE_minus <- gather(NH5_AE_minus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
```

###Using function to remove outliers and create a new dataframe in wide form and long form for plus
```{r}

noOut_NH5_AE_plus.list <- apply(NH5_AE_plus[,2:ncol(NH5_AE_plus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_NH5_AE_plus_sideways <- plyr::ldply(noOut_NH5_AE_plus.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_NH5_AE_plus <- data.frame(t(noOut_NH5_AE_plus_sideways[-1]))
colnames(noOut_NH5_AE_plus) <- noOut_NH5_AE_plus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
NH5_AE_plus_noOut_wideform <- cbind(NH5_AE_plus[1:nrow(noOut_NH5_AE_plus),1], noOut_NH5_AE_plus[,1:ncol(noOut_NH5_AE_plus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(NH5_AE_plus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_NH5_AE_plus <- gather(NH5_AE_plus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
```

##Make DF of medians of each after removal of outliers 
```{r}
#median DFs
#medianAE_FI<- as.data.frame(apply(noOut_NH5_AE_FI, 2, median, na.rm=TRUE))
medianAE_minus<- as.data.frame(apply(noOut_NH5_AE_minus, 2, median, na.rm=TRUE))
medianAE_plus<- as.data.frame(apply(noOut_NH5_AE_plus, 2, median, na.rm=TRUE))
#standard error DFs for minus and plus
SE_AE_minus <- std.error(noOut_NH5_AE_minus)
SE_AE_plus <- std.error(noOut_NH5_AE_plus)
#create one data frame that has medians and standard errors
AE_all_summary_table <- cbind.data.frame(medianAE_minus, SE_AE_minus, medianAE_plus, SE_AE_plus)

#rename column names
colnames(AE_all_summary_table) <- c("-2AP", "-2AP_SE","+2AP", "+2AP_SE")
#make FI
AE_all_summary_table$FI <- AE_all_summary_table$`+2AP` / AE_all_summary_table$`-2AP`
#divide values by NH5
AE_all_summary_table$corr_minus_2AP <- as.numeric(AE_all_summary_table$'-2AP')/AE_all_summary_table[c("NH5"),c("-2AP")]
AE_all_summary_table$corr_plus_2AP <- as.numeric(AE_all_summary_table$'+2AP')/AE_all_summary_table[c("NH5"),c("+2AP")]
AE_all_summary_table$corr_FI <- as.numeric(AE_all_summary_table$FI)/AE_all_summary_table[c("NH5"),c("FI")]

#the measurements should have 2 sig figs and the standards errors should have 1 sig fig
two_sf_cols <- c(1,3,5,6,7,8)
one_sf_cols <- c(2,4)
AE_all_summary_table[,two_sf_cols] <- signif(AE_all_summary_table[,two_sf_cols],2)
AE_all_summary_table[,one_sf_cols] <- signif(AE_all_summary_table[,one_sf_cols],1)

#write csv
write.csv(AE_all_summary_table,"~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/Anchor_exchange/AE_all_summary_table.csv", row.names = TRUE)

#remove the switches below the functional cutoff of 2.0
#find the FI median
AE_summary_table <- subset(AE_all_summary_table, FI>1.9)
#write csv
write.csv(AE_summary_table,"~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/Anchor_exchange/AE_summary_table.csv", row.names = TRUE)
#write excel
write.xlsx(AE_summary_table, '~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/Anchor_exchange/AE_summary_table.xlsx', rowNames = TRUE)
write.xlsx(AE_all_summary_table, '~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/Anchor_exchange/AE_all_summary_table.xlsx', rowNames = TRUE)
AE_all_summary_table_noGFP <- AE_all_summary_table[order(row.names(AE_all_summary_table)),]
AE_all_summary_table_noGFP <- AE_all_summary_table_noGFP[-(1),]
ggplot(AE_all_summary_table_noGFP, aes(x=`corr_minus_2AP`, y=`corr_FI`))+ geom_point(size=2, alpha=1, show.legend = FALSE)+
  labs(title="Anchor exchange Library Repression vs Activity", x="Repression", y="Fold Induction")+
  theme_minimal()
```
#binning by REP
```{r}
REPbinsAE <- AE_all_summary_table_noGFP
REPbadbinAE <- REPbinsAE %>% 
 filter(AE_all_summary_table_noGFP$`corr_minus_2AP` > 9)
REPbinAE <- rep(1,nrow(REPbadbinAE))
REPbadbinAE <- cbind(REPbadbinAE,REPbinAE)
REPmediumbinAE<- REPbinsAE %>% 
  filter(AE_all_summary_table_noGFP$`corr_minus_2AP` >=1.5 & AE_all_summary_table_noGFP$`corr_minus_2AP` <=9)
REPbinAE <- rep(6,nrow(REPmediumbinAE))
REPmediumbinAE <- cbind(REPmediumbinAE,REPbinAE)
REPgreatbinAE <- REPbinsAE %>% 
 filter(AE_all_summary_table_noGFP$`corr_minus_2AP` < 1.5)
REPbinAE <- rep(10,nrow(REPgreatbinAE))
REPgreatbinAE <- cbind(REPgreatbinAE,REPbinAE)
REPallbinsAE <- rbind(REPbadbinAE,REPmediumbinAE,REPgreatbinAE)
ggplot(REPallbinsAE, aes(x=`corr_minus_2AP`, y=`corr_FI`, color=REPbinAE))+ geom_point(size=2, alpha=1, show.legend = FALSE)+
  labs(title="AE Library Repression vs Activity", x="-2AP Fluorescence", y="Fold Induction")+
  theme_minimal()
tryREPallbinsAE <- REPallbinsAE[order(row.names(REPallbinsAE)),]
```

#binning by FI
```{r}
FIbinsAE <- AE_all_summary_table_noGFP
FIgoodbinAE <- FIbinsAE %>% 
 filter(AE_all_summary_table_noGFP$`corr_FI` >= 0.5)
FIbinAE <- rep(10,nrow(FIgoodbinAE))
FIgoodbinAE <- cbind(FIgoodbinAE,FIbinAE)
FIbadbinAE <- FIbinsAE %>% 
 filter(AE_all_summary_table_noGFP$`corr_FI` < 0.5)
FIbinAE <- rep(5,nrow(FIbadbinAE))
FIbadbinAE <- cbind(FIbadbinAE,FIbinAE)
FIallbinsAE <- rbind(FIbadbinAE,FIgoodbinAE)
ggplot(FIallbinsAE, aes(x=`corr_minus_2AP`, y=`corr_FI`, color=FIbinAE))+ geom_point(size=2, alpha=1, show.legend = FALSE)+
  labs(title="AE Library Repression vs Activity", x="-2AP Fluorescence", y="Fold Induction")+
  theme_minimal()
tryFIallbinsAE <- FIallbinsAE[order(row.names(FIallbinsAE)),]
```
#all bins together
```{r}
allinfoAE <- cbind(tryREPallbinsAE,tryFIallbinsAE[,"FIbinAE"])
colnames(allinfoAE) <- c("-2AP", "-2AP_SE","+2AP", "+2AP_SE", "FI", "corr_minus_2AP", "corr_plus_2AP", "corr_FI", "REPbin", "FIbin")
allinfoAE$binvalue <- allinfoAE$`REPbin` + allinfoAE$`FIbin`
allinfoAE$bin <- NA
allinfoAE$bin[allinfoAE$binvalue==20] <- "top"
allinfoAE$bin[allinfoAE$binvalue==16] <- "top"
allinfoAE$bin[allinfoAE$binvalue==11] <- "mid"
allinfoAE$bin[allinfoAE$binvalue==6] <- "bottom"
#name NH5
allinfoAE["NH5","bin"] <- "NH5"
ggplot(allinfoAE, aes(x=`corr_minus_2AP`, y=`corr_FI`, color=bin))+ geom_point(size=2, alpha=1, show.legend = FALSE)+
  labs(title="AE Library Repression vs Activity", x="-2AP Fluorescence", y="Fold Induction")+
  theme_minimal()
write.xlsx(allinfoAE, '~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/Anchor_exchange/AE_binning.xlsx', rowNames = TRUE)
```
#function to read basepairing from df
```{r}
basepairing <- function(base1,base2){
  pairing <- 0
  if(base1=="A"){
    if(base2=="U"){
      pairing <-1
    }
  }
  if(base1=="G"){
    if(base2=="U" | base2=="C"){
      pairing <-1
    }
  }
  if(base1=="U"){
    if(base2=="A" | base2=="G"){
      pairing <-1
    }
  }
  if(base1=="C"){
    if(base2=="G"){
      pairing <-1
    }
  }
 return(pairing)   
}   

weighted.basepairing <- function(base1,base2){
  pairing <- 0
  if(base1=="A"){
    if(base2=="U"){
      pairing <-1
    }
  }
  if(base1=="G"){
    if(base2=="U" ){
      pairing <-1
     } else if(base2=="C"){
        pairing <-2
    }
  }
  if(base1=="U"){
    if(base2=="A" | base2=="G"){
      pairing <-1
    }
  }
  if(base1=="C"){
    if(base2=="G"){
      pairing <-2
    }
  }
 return(pairing)   
 }  
 
```
#read in df with seq
```{r}
AE_summ_seq <- read.csv("~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/Anchor_exchange/AE_all_summary_table_seq.csv")
SEQ_AE <- AE_summ_seq[,c("X","P1.1","P1P4.1", "P4.2", "P4.3", "P4.4" )]
SEQ_AE <- filter(SEQ_AE,X!="GFPUV")
AE_all_summary_table <- filter(AE_all_summary_table, row.names(AE_all_summary_table)!="GFPUV")
SEQ_AE2 <- SEQ_AE
SEQ_AE1 <- SEQ_AE

```
#add score for number of bps in P1 and P4
```{r}
#df for score for P4 and score for P1
SEQ_AE <- SEQ_AE%>% 
  mutate(P1pair = mapply(basepairing,SEQ_AE[,2], SEQ_AE[,3]),P4_pairing = mapply(basepairing,SEQ_AE[,3],SEQ_AE[,6]) + mapply(basepairing,SEQ_AE[,4], SEQ_AE[,5]))
AE_all_summary_table <- rownames_to_column(AE_all_summary_table)
colnames(SEQ_AE)[1] <- c("rowname")
colnames(SEQ_AE1)[1] <- c("rowname")
colnames(SEQ_AE2)[1] <- c("rowname")
#df that has each bp separated, P1, P4-1 which is the bp distal to L4, and P4-2 which is proximal to L4
SEQ_AE1 <- SEQ_AE1%>% 
  mutate(P1pair = mapply(basepairing,SEQ_AE[,2], SEQ_AE[,3]),P4_pair1 = mapply(basepairing,SEQ_AE[,3],SEQ_AE[,6]), P4_pair2 = mapply(basepairing,SEQ_AE[,4], SEQ_AE[,5]))
AE_all_summary_table <- rownames_to_column(AE_all_summary_table)
colnames(SEQ_AE)[1] <- c("rowname")
#df that has GC as 2 and AU/GU as 1 and all others as 0
SEQ_AE2 <- SEQ_AE2%>% 
  mutate(P1pair = mapply(weighted.basepairing,SEQ_AE[,2], SEQ_AE[,3]),P4_pairing = mapply(weighted.basepairing,SEQ_AE[,3],SEQ_AE[,6]) + mapply(weighted.basepairing,SEQ_AE[,4], SEQ_AE[,5]))
```
##Plots
```{r}
#plain plot that has NH5 as another color 
is_NH5 <- rep(0,nrow(AE_all_summary_table))
is_NH5[1] <- 1
basicgraph <- cbind(AE_all_summary_table, as.factor(is_NH5))
ggplot(AE_all_summary_table, aes(x=`corr_minus_2AP`, y=`corr_FI`, color=is_NH5)) + geom_point(alpha=3,show.legend = FALSE)+
  labs(title="Anchor Exchange Fold Induction vs Repression", x="Repression", y="Fold Induction")+
  theme_minimal()
#graph based on whether there are 0,1, or 2 pairs in P4, shape based on whether or not there is a P1 extending pair
P4pairinggraph <- left_join(AE_all_summary_table,SEQ_AE,by="rowname")
P4pairinggraph$P4_pairing <- as.factor(P4pairinggraph$P4_pairing)
P4pairinggraph$P1pair <- as.factor(P4pairinggraph$P1pair)
ggplot(P4pairinggraph, aes(x=`corr_minus_2AP`, y=`corr_FI`, group=`P4_pairing`)) + 
  geom_point(size=3, alpha=1, aes(color=`P4_pairing`)) +
  labs(title="Anchor Exchange Fold Induction vs Repression", x="Repression", y="Fold Induction")+
  #scale_color_brewer(palette = "Paired",trans="reverse")+
  scale_color_manual(values=rev(brewer.pal(3,"Paired")))+
  theme_minimal()
  #theme(legend.position='none'))
#graphsbased on which P4 bp is present 
#is distal P4 bp present??
whichP4pairinggraph <- left_join(AE_all_summary_table,SEQ_AE1,by="rowname")
whichP4pairinggraph$P4_pair1 <- as.factor(whichP4pairinggraph$P4_pair1)
whichP4pairinggraph$P4_pair2 <- as.factor(whichP4pairinggraph$P4_pair2)
whichP4pairinggraph$P1pair <- as.factor(whichP4pairinggraph$P1pair)
ggplot(whichP4pairinggraph, aes(x=`corr_minus_2AP`, y=`corr_FI`)) + 
  geom_point(size=3, alpha=1, aes(color=`P4_pair1`)) +
  labs(title="Anchor Exchange Fold Induction vs Repression", x="Repression", y="Fold Induction")+
  #scale_color_brewer(palette = "Paired",trans="reverse")+
  scale_color_manual(values=rev(brewer.pal(3,"Paired")))+
  theme_minimal()
  #theme(legend.position='none'))
#is proximal P4 bp present??
ggplot(whichP4pairinggraph, aes(x=`corr_minus_2AP`, y=`corr_FI`)) + 
  geom_point(size=3, alpha=1, aes(color=`P4_pair2`)) +
  labs(title="Anchor Exchange Fold Induction vs Repression", x="Repression", y="Fold Induction")+
  #scale_color_brewer(palette = "Paired",trans="reverse")+
  scale_color_manual(values=rev(brewer.pal(3,"Paired")))+
  theme_minimal()
  #theme(legend.position='none'))
#does strength of bping in P4 matter??
strengthP4pairinggraph <- left_join(AE_all_summary_table,SEQ_AE2,by="rowname")
strengthP4pairinggraph$P1pair<- as.factor(strengthP4pairinggraph$P1pair)
strengthP4pairinggraph$P4_pairing <- as.factor(strengthP4pairinggraph$P4_pairing)
ggplot(strengthP4pairinggraph, aes(x=`corr_minus_2AP`, y=`corr_FI`)) + 
  geom_point(size=3, alpha=0.8, aes(color=`P4_pairing`)) +
  labs(title="Anchor Exchange Fold Induction vs Repression", x="Repression", y="Fold Induction")+
  #scale_color_brewer(palette = "Paired",trans="reverse")+
  scale_color_manual(values=rev(brewer.pal(5,"Paired")))+
  theme_minimal()
  #theme(legend.position='none'))

P4pairinggraph <- left_join(AE_all_summary_table,SEQ_AE,by="rowname")
P4pairinggraph$P4_pairing <- as.factor(P4pairinggraph$P4_pairing)
P4pairinggraph$P1pair <- as.factor(P4pairinggraph$P1pair)
ggplot(P4pairinggraph, aes(x=`corr_minus_2AP`, y=`corr_FI`, group=`P4_pairing`)) + 
  geom_point(size=3, alpha=1, aes(color=`P1pair`)) +
  labs(title="Anchor Exchange Fold Induction vs Repression", x="Repression", y="Fold Induction")+
  #scale_color_brewer(palette = "Paired",trans="reverse")+
  scale_color_manual(values=rev(brewer.pal(3,"Paired")))+
  theme_minimal()
  #theme(legend.position='none'))

#Does it matter which bp is formed in P4?? Made column to color graph for which P4 bp is formed, those that form 0 bps or 2 bps are treated equal in this to isolate single bps, distal is the pair distal to L4, proximal is P4 pair proximal to L4
whichP4pairinggraph <- whichP4pairinggraph%>% 
  mutate(whichP4 = ifelse(P4_pair1==1 & P4_pair2==0,"distal_P4",ifelse(P4_pair1==0 & P4_pair2==1,"proximal_P4","not_1_bp")))
whichP4pairinggraph$whichP4 <- as.factor(whichP4pairinggraph$whichP4)
ggplot(whichP4pairinggraph, aes(x=`corr_minus_2AP`, y=`corr_FI`)) + 
  geom_point(size=3, alpha=1, aes(color=`whichP4`)) +
  labs(title="Anchor Exchange Fold Induction vs Repression", x="Repression", y="Fold Induction")+
  #scale_color_brewer(palette = "Paired",trans="reverse")+
  #scale_color_manual(values=rev(brewer.pal(3,"Paired")))+
  theme_minimal()+
  #scale_color_manual(values= c("#78B7C5", "#EBCC2A", "#F21A00"))
  #scale_color_manual(values= c("#3B9AB2", "#EBCC2A", "#F21A00"))
  scale_color_manual(values= c( "#F21A00","dark gray", "#3B9AB2"))
  #theme(legend.position='none'))
#separate plots based on the base identity of P1
ggplot(filter(whichP4pairinggraph, P1.1 == "A"), aes(x=`corr_minus_2AP`, y=`corr_FI`)) + 
  geom_point(size=3, alpha=1) +
  labs(title="Anchor Exchange Fold Induction vs Repression When P1=A", x="Repression", y="Fold Induction")+
  #scale_color_brewer(palette = "Paired",trans="reverse")+
  #scale_color_manual(values=rev(brewer.pal(3,"Paired")))+
  theme_minimal()+
  xlim(0,20)+
  ylim(0,2)+
  #scale_color_manual(values= c("#78B7C5", "#EBCC2A", "#F21A00"))
  #scale_color_manual(values= c("#3B9AB2", "#EBCC2A", "#F21A00"))
  scale_color_manual(values= c( "#F21A00","dark gray", "#3B9AB2"))
  #theme(legend.position='none'))
ggplot(filter(whichP4pairinggraph, P1.1 == "C"), aes(x=`corr_minus_2AP`, y=`corr_FI`)) + 
  geom_point(size=3, alpha=1) +
  labs(title="Anchor Exchange Fold Induction vs Repression When P1=C", x="Repression", y="Fold Induction")+
  #scale_color_brewer(palette = "Paired",trans="reverse")+
  #scale_color_manual(values=rev(brewer.pal(3,"Paired")))+
  theme_minimal()+
  #scale_color_manual(values= c("#78B7C5", "#EBCC2A", "#F21A00"))
  #scale_color_manual(values= c("#3B9AB2", "#EBCC2A", "#F21A00"))
  scale_color_manual(values= c( "#F21A00","dark gray", "#3B9AB2"))
  #theme(legend.position='none'))
ggplot(filter(whichP4pairinggraph, P1.1 == "G"), aes(x=`corr_minus_2AP`, y=`corr_FI`)) + 
  geom_point(size=3, alpha=1) +
  labs(title="Anchor Exchange Fold Induction vs Repression When P1=G", x="Repression", y="Fold Induction")+
  #scale_color_brewer(palette = "Paired",trans="reverse")+
  #scale_color_manual(values=rev(brewer.pal(3,"Paired")))+
  theme_minimal()+
  #scale_color_manual(values= c("#78B7C5", "#EBCC2A", "#F21A00"))
  #scale_color_manual(values= c("#3B9AB2", "#EBCC2A", "#F21A00"))
  scale_color_manual(values= c( "#F21A00","dark gray", "#3B9AB2"))
  #theme(legend.position='none'))
ggplot(filter(whichP4pairinggraph, P1.1 == "U"), aes(x=`corr_minus_2AP`, y=`corr_FI`)) + 
  geom_point(size=3, alpha=1) +
  labs(title="Anchor Exchange Fold Induction vs Repression When P1=U", x="Repression", y="Fold Induction")+
  #scale_color_brewer(palette = "Paired",trans="reverse")+
  #scale_color_manual(values=rev(brewer.pal(3,"Paired")))+
  theme_minimal()+
  #scale_color_manual(values= c("#78B7C5", "#EBCC2A", "#F21A00"))
  #scale_color_manual(values= c("#3B9AB2", "#EBCC2A", "#F21A00"))
  scale_color_manual(values= c( "#F21A00","dark gray", "#3B9AB2"))
  #theme(legend.position='none'))
#all 4 at once
ggplot(whichP4pairinggraph, aes(x=`corr_minus_2AP`, y=`corr_FI`)) + 
  geom_point(size=3, alpha=1,color="blue") +
  labs(title="Anchor Exchange Fold Induction vs Repression As P1 varies", x="Repression", y="Fold Induction")+
  theme_minimal()+
  facet_wrap(vars(P1.1))
```

```{r}

##P1 color plot
ggplot(sillygraph, aes(x=`corr_minus_2AP`, y=`corr_FI`, group=`P4_pairing`)) + 
  geom_point(size=3, alpha=0.75, aes(color=`P1pair`)) +
  labs(title="Anchor Exchange Fold Induction vs Repression", x="Repression", y="Fold Induction")+
  #scale_color_brewer(palette = "Paired",trans="reverse")+
  scale_color_manual(values=rev(brewer.pal(3,"Paired")))+
  theme_minimal()
  #theme(legend.position='none'))

##P4 color plot
ggplot(sillygraph, aes(x=`corr_minus_2AP`, y=`corr_FI`, group=`P4_pairing`)) + 
  geom_point(size=3, alpha=0.75, aes(color=`P4_pairing`)) +
  labs(title="Anchor Exchange Fold Induction vs Repression", x="Repression", y="Fold Induction")+
  #scale_color_brewer(palette = "Paired",trans="reverse")+
  scale_color_manual(values=rev(brewer.pal(3,"Paired")))+
  theme_minimal()
  #theme(legend.position='none'))
```

