---
title: "L4_analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(RColorBrewer)
library(plyr)
library(dplyr)
library(tibble)
library(ggsignif)
library(plotrix)
library(openxlsx)
library(viridis)
```

## Creating the function to remove outliers based on Interquartile range
```{r}
removeOut <- function(distribution, bounds=1.5, na.rm = TRUE) {
  #finds the 25th and 75th percentile of the distribution, excluding NAs
  Q <- quantile(distribution, probs=c(.25,.75), na.rm=TRUE) 
 #difference between 75th and 25th quartile=interquartile range
 iqr <- IQR(distribution, na.rm=TRUE)
 #upper bound of the values is 75th quartile + 1.5*(interquartile range) note: that 1.5 can be varied by changing bound
 upperBound <- Q[2]+bounds*iqr
 #lower bound of the values is 25th quartile - 1.5*(interquartile range)
 lowerBound <- Q[1]-bounds*iqr
 #take out the values that are outside of the established bounds note: that 1.5 can be varied by changing bound
 eliminated <- c(distribution < lowerBound | distribution > upperBound)
 #index values within the upper and lower bounds
 noteliminatedindex <- which(distribution > lowerBound & distribution < upperBound)
 noteliminatedvalues <- distribution[noteliminatedindex]
 #return those values which will be included that are between the upper and lower bounds 
 return(noteliminatedvalues)
 
}
```

Read in csv.
Formatting of csv: each column is a different colony, first row is the name of the colony, each column is then filled with the background fluorescence (PBR322) corrected fluorescence values (there should be three values per biological replicate, each of which are the technical replicates). Far right column is the condition of the growth, either - or + ligand. For this particular switch, -L=repression and +L=expression. For those colonies that have less raw data values than the controls, the empty cells are filled with an integer that can be removed easily (-2) to be turned into NA when read into R. 
This is for repression and expression, fold induction will not be included because it will be calculated based on the included repression and expression values.
```{r}
#read in csv, these values will include colonies below the function cut off which has been defined here as 2.0
L4_all <- read.csv("~/Downloads/Expression_platform_project/expression_platform_files/L4_data.csv", header=T)
empty_columns <- sapply(L4_all, function(x) all(is.na(x) | x==""))
L4_all <- L4_all[,!empty_columns]
L4_all[L4_all == -2] <-NA
#MINUS 2AP DF
L4_minus1 <- L4_all[L4_all == "(-)",]
#remove rows that were only NA and delete DF that was filled with NA only columns
L4_minus <- L4_minus1[rowSums(is.na(L4_minus1)) !=ncol(L4_minus1),]
rm(L4_minus1)
#PLUS 2AP DF
L4_plus1 <- L4_all[L4_all == "(+)",]
L4_plus <- L4_plus1[rowSums(is.na(L4_plus1)) !=ncol(L4_plus1),]
rm(L4_plus1)
```

###Using function to remove outliers and create a new dataframe in wide form and long form
```{r}
#apply function to remove outliers for minus
noOut_L4_minus.list <- apply(L4_minus[,2:ncol(L4_minus)], 2, removeOut, na.rm=TRUE)
noOut_L4_minus_sideways <- plyr::ldply(noOut_L4_minus.list, rbind)
noOut_L4_minus <- data.frame(t(noOut_L4_minus_sideways[-1]))
colnames(noOut_L4_minus) <- noOut_L4_minus_sideways[,1]
#convert to long form
L4_minus_noOut_wideform <- cbind(L4_minus[1:nrow(noOut_L4_minus),1],noOut_L4_minus[,1:ncol(noOut_L4_minus)])
colnames(L4_minus_noOut_wideform)[1] <- "COLONIES"
long_L4_minus <- gather(L4_minus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
#apply function to remove outliers for plus
noOut_L4_plus.list <- apply(L4_plus[,2:ncol(L4_plus)], 2, removeOut, na.rm=TRUE)
noOut_L4_plus_sideways <- plyr::ldply(noOut_L4_plus.list, rbind)
noOut_L4_plus <- data.frame(t(noOut_L4_plus_sideways[-1]))
colnames(noOut_L4_plus) <- noOut_L4_plus_sideways[,1]
#convert to long form
L4_plus_noOut_wideform <- cbind(L4_plus[1:nrow(noOut_L4_plus),1],noOut_L4_plus[,1:ncol(noOut_L4_plus)])
colnames(L4_plus_noOut_wideform)[1] <- "COLONIES"
long_L4_plus <- gather(L4_plus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
```

##Make DF of medians of each after removal of outliers 
```{r}
#median DFs
medianL4_minus<- as.data.frame(apply(noOut_L4_minus, 2, median, na.rm=TRUE))
medianL4_plus<- as.data.frame(apply(noOut_L4_plus, 2, median, na.rm=TRUE))
#standard error DFs for minus and plus
SE_L4_minus <- std.error(noOut_L4_minus)
SE_L4_plus <- std.error(noOut_L4_plus)
#create one data frame that has medians and standard errors
L4_all_summary_table <- cbind.data.frame(medianL4_minus, SE_L4_minus, medianL4_plus, SE_L4_plus)
colnames(L4_all_summary_table) <- c("-2AP", "-2AP_SE","+2AP", "+2AP_SE")
#make FI
L4_all_summary_table$FI <- L4_all_summary_table$`+2AP` / L4_all_summary_table$`-2AP`
#divide values by P4A
L4_all_summary_table$corr_minus_2AP <- as.numeric(L4_all_summary_table$'-2AP')/L4_all_summary_table[c("P4A"),c("-2AP")]
L4_all_summary_table$corr_plus_2AP <- as.numeric(L4_all_summary_table$'+2AP')/L4_all_summary_table[c("P4A"),c("+2AP")]
L4_all_summary_table$corr_FI <- as.numeric(L4_all_summary_table$FI)/L4_all_summary_table[c("P4A"),c("FI")]
#the measurements should have 2 sig figs and the standards errors should have 1 sig fig
two_sf_cols <- c(1,3,5,6,7,8)
one_sf_cols <- c(2,4)
L4_all_summary_table[,two_sf_cols] <- signif(L4_all_summary_table[,two_sf_cols],2)
L4_all_summary_table[,one_sf_cols] <- signif(L4_all_summary_table[,one_sf_cols],1)
#write csv
write.csv(L4_all_summary_table,"~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/L4/L4_all_summary_table.csv", row.names = TRUE)
#remove the switches below the functional cutoff of 2.0
#find the FI median
L4_summary_table <- subset(L4_all_summary_table, FI>1.9)
#write csv
write.csv(L4_summary_table,"~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/L4/L4_summary_table.csv", row.names = TRUE)
#write excel
write.xlsx(L4_summary_table, '~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/L4/L4_summary_table.xlsx', rowNames = TRUE)
write.xlsx(L4_all_summary_table, '~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/L4/L4_all_summary_table.xlsx', rowNames = TRUE)
#remove GFP from graph
L4_all_summary_table_noGFP <- L4_all_summary_table[order(row.names(L4_all_summary_table)),]
L4_all_summary_table_noGFP <- L4_all_summary_table_noGFP[-(1),]
ggplot(L4_all_summary_table_noGFP, aes(x=`corr_minus_2AP`, y=`corr_FI`))+ geom_point(size=2, alpha=1, show.legend = FALSE)+
  labs(title="L4 Library Repression vs Activity", x="-2AP Fluorescence", y="Fold Induction")+
  theme_minimal()
```
