---
title: "n6_p4stem_analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(RColorBrewer)
library(plyr)
library(dplyr)
library(tibble)
library(ggsignif)
library(plotrix)
library(openxlsx)
library(data.table)
```

## Creating the function to remove outliers based on Interquartile range
```{r}
removeOut <- function(distribution, bounds=1.5, na.rm = TRUE) {
 Q <- quantile(distribution, probs=c(.25,.75), na.rm=TRUE) 
 #finds the 25th and 75th percentile of the distribution, excluding NAs
 iqr <- IQR(distribution, na.rm=TRUE)
 #difference between 75th and 25th quartile=interquartile range
 upperBound <- Q[2]+bounds*iqr
 #upper bound of the values is 75th quartile + 1.5*(interquartile range) note: that 1.5 can be varied by changing bound
 lowerBound <- Q[1]-bounds*iqr
 #upper bound of the values is 25th quartile - 1.5*(interquartile range)
 eliminated <- c(distribution < lowerBound | distribution > upperBound)
 #take out the values that are outside of the established bounds note: that 1.5 can be varied by changing bound
 noteliminatedindex <- which(distribution > lowerBound & distribution < upperBound)
 #index values within the upper and lower bounds
 noteliminatedvalues <- distribution[noteliminatedindex]
 return(noteliminatedvalues)
 #return those values which will be included that are between the upper and lower bounds 
}
```

Read in csv.
Formatting of csv: each column is a different colony, first row is the name of the colony, each column is then filled with the background fluorescence (PBR322) corrected fluorescence values (there should be three values per biological replicate, each of which are the technical replicates). Far right column is the condition of the growth, either - or + ligand. For this particular switch, -L=repression and +L=expression. For those colonies that have less raw data values than the controls, the empty cells are filled with an integer that can be removed easily (-2) to be turned into NA when read into R. 
This is for repression and expression, fold induction will not be included because it will be calculated based on the included repression and expression values.
```{r}
#read in csv, these values will include colonies below the function cut off which has been defined here as 2.0
NH5_n6p4stemall <- read.csv("~/Downloads/Expression_platform_project/expression_platform_files/N6_P4stem_data (4).csv", header=T)
empty_columns <- sapply(NH5_n6p4stemall, function(x) all(is.na(x) | x==""))
NH5_n6p4stemall <- NH5_n6p4stemall[,!empty_columns]
NH5_n6p4stemall[NH5_n6p4stemall == -2] <-NA
#MINUS 2AP DF
NH5_n6p4stem_minus1 <- NH5_n6p4stemall[NH5_n6p4stemall == "(-)",]
NH5_n6p4stem_minus <- NH5_n6p4stem_minus1[rowSums(is.na(NH5_n6p4stem_minus1)) !=ncol(NH5_n6p4stem_minus1),]
rm(NH5_n6p4stem_minus1)
#remove rows that were only NA and delete DF that was filled with NA only columns
#PLUS 2AP DF
NH5_n6p4stem_plus1 <- NH5_n6p4stemall[NH5_n6p4stemall == "(+)",]
NH5_n6p4stem_plus <- NH5_n6p4stem_plus1[rowSums(is.na(NH5_n6p4stem_plus1)) !=ncol(NH5_n6p4stem_plus1),]
rm(NH5_n6p4stem_plus1)
#remove rows that were only NA 
```

###Using function to remove outliers and create a new dataframe in wide form and long form for minus
```{r}
noOut_NH5_n6p4stem_minus.list <- apply(NH5_n6p4stem_minus[,2:ncol(NH5_n6p4stem_minus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_NH5_n6p4stem_minus_sideways <- plyr::ldply(noOut_NH5_n6p4stem_minus.list, rbind)
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_NH5_n6p4stem_minus <- data.frame(t(noOut_NH5_n6p4stem_minus_sideways[-1]))
colnames(noOut_NH5_n6p4stem_minus) <- noOut_NH5_n6p4stem_minus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
NH5_n6p4stem_minus_noOut_wideform <- cbind(NH5_n6p4stem_minus[1:nrow(noOut_NH5_n6p4stem_minus),1], noOut_NH5_n6p4stem_minus[,1:ncol(noOut_NH5_n6p4stem_minus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(NH5_n6p4stem_minus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_NH5_n6p4stem_minus <- gather(NH5_n6p4stem_minus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
```

###Using function to remove outliers and create a new dataframe in wide form and long form for plus
```{r}

noOut_NH5_n6p4stem_plus.list <- apply(NH5_n6p4stem_plus[,2:ncol(NH5_n6p4stem_plus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_NH5_n6p4stem_plus_sideways <- plyr::ldply(noOut_NH5_n6p4stem_plus.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_NH5_n6p4stem_plus <- data.frame(t(noOut_NH5_n6p4stem_plus_sideways[-1]))
colnames(noOut_NH5_n6p4stem_plus) <- noOut_NH5_n6p4stem_plus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
NH5_n6p4stem_plus_noOut_wideform <- cbind(NH5_n6p4stem_plus[1:nrow(noOut_NH5_n6p4stem_plus),1], noOut_NH5_n6p4stem_plus[,1:ncol(noOut_NH5_n6p4stem_plus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(NH5_n6p4stem_plus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_NH5_n6p4stem_plus <- gather(NH5_n6p4stem_plus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
```



##Make DF of medians of each after removal of outliers 
```{r}
#median DFs
#mediann6p4stem_FI<- as.data.frame(apply(noOut_NH5_n6p4stem_FI, 2, median, na.rm=TRUE))
mediann6p4stem_minus<- as.data.frame(apply(noOut_NH5_n6p4stem_minus, 2, median, na.rm=TRUE))
##Remove negative values
#Those values that have negative repression values are because they repress so well that they are similar to the pbr322 values...Negative values will be replaced with a repression value of 100 which is lower than all other measured values because that is outside of the accuracy of the measurements of this assay
positive_mediann6p4stem_minus <- mediann6p4stem_minus
positive_mediann6p4stem_minus[positive_mediann6p4stem_minus < 0] <- 50
mediann6p4stem_plus<- as.data.frame(apply(noOut_NH5_n6p4stem_plus, 2, median, na.rm=TRUE))
#standard error DFs for minus and plus
SE_n6p4stem_minus <- std.error(noOut_NH5_n6p4stem_minus)
SE_n6p4stem_plus <- std.error(noOut_NH5_n6p4stem_plus)
#create one data frame that has medians and standard errors
n6p4stem_all_summary_table <- cbind.data.frame(positive_mediann6p4stem_minus, SE_n6p4stem_minus, mediann6p4stem_plus, SE_n6p4stem_plus)

#rename column names
colnames(n6p4stem_all_summary_table) <- c("-2AP", "-2AP_SE","+2AP", "+2AP_SE")
#make FI
n6p4stem_all_summary_table$FI <- n6p4stem_all_summary_table$`+2AP` / n6p4stem_all_summary_table$`-2AP`
#divide values by NH5
n6p4stem_all_summary_table$corr_minus_2AP <- as.numeric(n6p4stem_all_summary_table$'-2AP')/n6p4stem_all_summary_table[c("NH5"),c("-2AP")]
n6p4stem_all_summary_table$corr_plus_2AP <- as.numeric(n6p4stem_all_summary_table$'+2AP')/n6p4stem_all_summary_table[c("NH5"),c("+2AP")]
n6p4stem_all_summary_table$corr_FI <- as.numeric(n6p4stem_all_summary_table$FI)/n6p4stem_all_summary_table[c("NH5"),c("FI")]

#the measurements should have 2 sig figs and the standards errors should have 1 sig fig
two_sf_cols <- c(1,3,5,6,7,8)
one_sf_cols <- c(2,4)
n6p4stem_all_summary_table[,two_sf_cols] <- signif(n6p4stem_all_summary_table[,two_sf_cols],2)
n6p4stem_all_summary_table[,one_sf_cols] <- signif(n6p4stem_all_summary_table[,one_sf_cols],1)

#write csv
write.csv(n6p4stem_all_summary_table,"~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/n6_P4stem/n6p4stem_all_summary_table.csv", row.names = TRUE)

#remove the switches below the functional cutoff of 2.0
#find the FI median
n6p4stem_summary_table <- subset(n6p4stem_all_summary_table, FI>1.9)
#write csv
write.csv(n6p4stem_summary_table,"~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/n6_P4stem/n6p4stem_summary_table.csv", row.names = TRUE)
#write excel
write.xlsx(n6p4stem_summary_table, '~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/n6_P4stem/n6p4stem_summary_table.xlsx', rowNames = TRUE)
write.xlsx(n6p4stem_all_summary_table, '~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/n6_P4stem/n6p4stem_all_summary_table.xlsx', rowNames = TRUE)
#REMOVE GFPUV
N6P4STEM_all_summary_table_noGFP <- n6p4stem_all_summary_table[order(row.names(n6p4stem_all_summary_table)),]
N6P4STEM_all_summary_table_noGFP <- N6P4STEM_all_summary_table_noGFP[-(1),]

ggplot(N6P4STEM_all_summary_table_noGFP, aes(x=`corr_minus_2AP`, y=`corr_FI`))+ geom_point(size=2, alpha=1, show.legend = FALSE)+
  labs(title="N=6 P4 Stem Library Repression vs Activity", x="-2AP Fluorescence", y="Fold Induction")+
  theme_minimal()


```

```{r}
#make a df without removing outliers to compare the values before and after outlier removal 
raw_minus <- NH5_n6p4stem_minus
raw_minus_med <- as.data.frame(apply(raw_minus[,2:ncol(raw_minus)], 2, median, na.rm=TRUE))
raw_plus <- NH5_n6p4stem_plus
raw_plus_med <- as.data.frame(apply(raw_plus[,2:ncol(raw_plus)], 2, median, na.rm=TRUE))
raw_minus_SE <- std.error(raw_minus[,2:ncol(raw_minus)])
raw_plus_SE <- std.error(raw_plus[,2:ncol(raw_plus)])
raw_summary_table <- cbind(raw_minus_med, raw_minus_SE, raw_plus_med, raw_plus_SE)
colnames(raw_summary_table) <- c("-2AP", "-2AP_SE","+2AP", "+2AP_SE")
#make FI
raw_summary_table$FI <- raw_summary_table$`+2AP` / raw_summary_table$`-2AP`
two_sf_cols <- c(1,3,5)
one_sf_cols <- c(2,4)
raw_summary_table[,two_sf_cols] <- signif(raw_summary_table[,two_sf_cols],2)
raw_summary_table[,one_sf_cols] <- signif(raw_summary_table[,one_sf_cols],1)
write.xlsx(raw_summary_table, '~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/n6_P4stem/raw_summary_table.xlsx', rowNames = TRUE)
```

##Analysis of colonies that were close to pbr322 

R10N.1B1 <- as.data.frame(noOut_NH5_n6p4stem_minus$R10N.1B1)
name <- rep("R10N.1B1",nrow(R10N.1B1))
R10N.1B1 <- cbind(R10N.1B1,name)
R10N.1B1 <- na.omit(R10N.1B1)
colnames(R10N.1B1) <- c("-2AP fluorescence","colony")
R3N.2D4 <- as.data.frame(noOut_NH5_n6p4stem_minus$R3N.2D4)
name <- rep("R3N.2D4",nrow(R3N.2D4))
R3N.2D4 <- cbind(R3N.2D4,name)
R3N.2D4 <- na.omit(R3N.2D4)
colnames(R3N.2D4) <- c("-2AP fluorescence","colony")
neg_df_rep <- rbind(R10N.1B1, R3N.2D4)

ggplot(neg_df_rep, aes(x=`colony`,`-2AP fluorescence`, color=`colony`)) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  geom_boxplot() +
  theme(legend.position="none")

group_by(neg_df_rep, colony) %>%
  summarise(
    count = n(),
    mean = mean(`-2AP fluorescence`, na.rm = TRUE),
    sd = sd(`-2AP fluorescence`, na.rm = TRUE)
  )
with(neg_df_rep, shapiro.test(`-2AP fluorescence`[colony == "R10N.1B1"]))
with(neg_df_rep, shapiro.test(`-2AP fluorescence`[colony == "R3N.2D4"]))
#both p values are greater than 0.05, meaning they are not significantly different from the normal distribution
res.ftest <- var.test(`-2AP fluorescence` ~ colony, data = neg_df_rep)
res.ftest
#p values is less than 0.05 so we cannot use standard t-test

#binning by REP
```{r}
REPbins <- N6P4STEM_all_summary_table_noGFP
REPbadbin <- REPbins %>% 
 filter(N6P4STEM_all_summary_table_noGFP$`corr_minus_2AP` > 9)
REPbin <- rep(1,nrow(REPbadbin))
REPbadbin <- cbind(REPbadbin,REPbin)
REPmediumbin<- REPbins %>% 
  filter(N6P4STEM_all_summary_table_noGFP$`corr_minus_2AP` >=1.5 & N6P4STEM_all_summary_table_noGFP$`corr_minus_2AP` <=9)
REPbin <- rep(6,nrow(REPmediumbin))
REPmediumbin <- cbind(REPmediumbin,REPbin)
REPgreatbin <- REPbins %>% 
 filter(N6P4STEM_all_summary_table_noGFP$`corr_minus_2AP` < 1.5)
REPbin <- rep(10,nrow(REPgreatbin))
REPgreatbin <- cbind(REPgreatbin,REPbin)
REPallbins <- rbind(REPbadbin,REPmediumbin,REPgreatbin)
ggplot(REPallbins, aes(x=`corr_minus_2AP`, y=`corr_FI`, color=REPbin))+ geom_point(size=2, alpha=1, show.legend = FALSE)+
  labs(title="N=6 P4 Stem Library Repression vs Activity", x="-2AP Fluorescence", y="Fold Induction")+
  theme_minimal()
tryREPallbins <- REPallbins[order(row.names(REPallbins)),]
```

#binning by FI
```{r}
FIbins <- N6P4STEM_all_summary_table_noGFP
FIgoodbin <- FIbins %>% 
 filter(N6P4STEM_all_summary_table_noGFP$`corr_FI` >= 0.5)
FIbin <- rep(10,nrow(FIgoodbin))
FIgoodbin <- cbind(FIgoodbin,FIbin)
FIbadbin <- FIbins %>% 
 filter(N6P4STEM_all_summary_table_noGFP$`corr_FI` < 0.5)
FIbin <- rep(5,nrow(FIbadbin))
FIbadbin <- cbind(FIbadbin,FIbin)
FIallbins <- rbind(FIbadbin,FIgoodbin)
ggplot(FIallbins, aes(x=`corr_minus_2AP`, y=`corr_FI`, color=FIbin))+ geom_point(size=2, alpha=1, show.legend = FALSE)+
  labs(title="N=6 P4 Stem Library Repression vs Activity", x="-2AP Fluorescence", y="Fold Induction")+
  theme_minimal()
tryFIallbins <- FIallbins[order(row.names(FIallbins)),]
```
#all bins together
```{r}
allinformation <- cbind(tryREPallbins,tryFIallbins[,"FIbin"])
colnames(allinformation) <- c("-2AP", "-2AP_SE","+2AP", "+2AP_SE", "FI", "corr_minus_2AP", "corr_plus_2AP", "corr_FI", "REPbinP3", "FIbinP3")
allinformation$binvalue <- allinformation$`REPbin` + allinformation$`FIbin`
allinformation$bin <- NA
allinformation$bin[allinformation$binvalue==20] <- "top"
allinformation$bin[allinformation$binvalue==16] <- "top"
allinformation$bin[allinformation$binvalue==11] <- "mid"
allinformation$bin[allinformation$binvalue==6] <- "bottom"
#name NH5
allinformation["NH5","bin"] <- "NH5"
ggplot(allinformation, aes(x=`corr_minus_2AP`, y=`corr_FI`, color=bin))+ geom_point(size=2, alpha=1, show.legend = FALSE)+
  labs(title="N=6 P4 Stem Library Repression vs Activity", x="-2AP Fluorescence", y="Fold Induction")+
  theme_minimal()
write.xlsx(allinformation, '~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/n6_P4stem/n6p4stem_binning.xlsx', rowNames = TRUE)
write.csv(allinformation, '~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/n6_P4stem/n6p4stem_binning.csv', row.names = TRUE)
```


```{r}
NOTCORRn6p4stem <- read.csv("~/Downloads/Expression_platform_project/expression_platform_files/notpbrcorr_N6_P4stem_data.csv", header=T)
empty_columns <- sapply(NOTCORRn6p4stem, function(x) all(is.na(x) | x==""))
NOTCORRn6p4stem <- NOTCORRn6p4stem[,!empty_columns]
NOTCORRn6p4stem[NOTCORRn6p4stem == -2] <-NA
#MINUS 2AP DF
NOTCORRn6p4stem_minus1 <- NOTCORRn6p4stem[NOTCORRn6p4stem == "(-)",]
NOTCORRn6p4stem_minus <- NOTCORRn6p4stem_minus1[rowSums(is.na(NOTCORRn6p4stem_minus1)) !=ncol(NOTCORRn6p4stem_minus1),]
rm(NOTCORRn6p4stem_minus1)
#remove rows that were only NA and delete DF that was filled with NA only columns
#PLUS 2AP DF
NOTCORRn6p4stem_plus1 <- NOTCORRn6p4stem[NOTCORRn6p4stem == "(+)",]
NOTCORRn6p4stem_plus <- NOTCORRn6p4stem_plus1[rowSums(is.na(NOTCORRn6p4stem_plus1)) !=ncol(NOTCORRn6p4stem_plus1),]
rm(NOTCORRn6p4stem_plus1)
#remove rows that were only NA 

noOut_NOTCORRn6p4stem_minus.list <- apply(NOTCORRn6p4stem_minus[,2:ncol(NOTCORRn6p4stem_minus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_NOTCORRn6p4stem_minus_sideways <- plyr::ldply(noOut_NOTCORRn6p4stem_minus.list, rbind)
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_NOTCORRn6p4stem_minus <- data.frame(t(noOut_NOTCORRn6p4stem_minus_sideways[-1]))
colnames(noOut_NOTCORRn6p4stem_minus) <- noOut_NOTCORRn6p4stem_minus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
NOTCORRn6p4stem_minus_noOut_wideform <- cbind(NOTCORRn6p4stem_minus[1:nrow(noOut_NOTCORRn6p4stem_minus),1], noOut_NOTCORRn6p4stem_minus[,1:ncol(noOut_NOTCORRn6p4stem_minus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(NOTCORRn6p4stem_minus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_NOTCORRn6p4stem_minus <- gather(NOTCORRn6p4stem_minus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
```
##Analysis of colonies that were close to pbr322 
```{r}
PBR <- as.data.frame(noOut_NOTCORRn6p4stem_minus$PBR)
name <- rep("PBR",nrow(PBR))
PBR <- cbind(PBR,name)
PBR <- na.omit(PBR)
colnames(PBR) <- c("-2AP fluorescence","colony")
R10N.1B1 <- as.data.frame(noOut_NOTCORRn6p4stem_minus$R10N.1B1)
name <- rep("R10N.1B1",nrow(R10N.1B1))
R10N.1B1 <- cbind(R10N.1B1,name)
R10N.1B1 <- na.omit(R10N.1B1)
colnames(R10N.1B1) <- c("-2AP fluorescence","colony")
R3N.2D4 <- as.data.frame(noOut_NOTCORRn6p4stem_minus$R3N.2D4)
name <- rep("R3N.2D4",nrow(R3N.2D4))
R3N.2D4 <- cbind(R3N.2D4,name)
R3N.2D4 <- na.omit(R3N.2D4)
colnames(R3N.2D4) <- c("-2AP fluorescence","colony")
neg_df_rep <- rbind(PBR,R10N.1B1, R3N.2D4)

ggplot(neg_df_rep, aes(x=`colony`,`-2AP fluorescence`, color=`colony`)) + 
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  geom_boxplot() +
  theme(legend.position="none")

group_by(neg_df_rep, colony) %>%
  summarise(
    count = n(),
    mean = mean(`-2AP fluorescence`, na.rm = TRUE),
    sd = sd(`-2AP fluorescence`, na.rm = TRUE)
  )

R10b1 <- neg_df_rep %>%
  filter(colony != "R3N.2D4")

  t.test(`-2AP fluorescence` ~ colony, data=R10b1)
R2d4 <- neg_df_rep %>%
  filter(colony != "R10N.1B1")

  t.test(`-2AP fluorescence` ~ colony, data=R2d4)
```