---
title: "P3_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(RColorBrewer)
library(plyr)
library(dplyr)
library(tibble)
library(ggsignif)
library(plotrix)
library(openxlsx)
```

## Creating the function to remove outliers based on Interquartile range

```{r}
removeOut <- function(distribution, bounds=1.5, na.rm = TRUE) {
 Q <- quantile(distribution, probs=c(.25,.75), na.rm=TRUE) 
 #finds the 25th and 75th percentile of the distribution, excluding NAs
 iqr <- IQR(distribution, na.rm=TRUE)
 #difference between 75th and 25th quartile=interquartile range
 upperBound <- Q[2]+bounds*iqr
 #upper bound of the values is 75th quartile + 1.5*(interquartile range) note: that 1.5 can be varied by changing bound
 lowerBound <- Q[1]-bounds*iqr
 #upper bound of the values is 25th quartile - 1.5*(interquartile range)
 eliminated <- c(distribution < lowerBound | distribution > upperBound)
 #take out the values that are outside of the established bounds note: that 1.5 can be varied by changing bound
 noteliminatedindex <- which(distribution > lowerBound & distribution < upperBound)
 #index values within the upper and lower bounds
 noteliminatedvalues <- distribution[noteliminatedindex]
 return(noteliminatedvalues)
 #return those values which will be included that are between the upper and lower bounds 
}
```

Read in csv.
Formatting of csv: each column is a different colony, first row is the name of the colony, each column is then filled with the background fluorescence (PBR322) corrected fluorescence values (there should be three values per biological replicate, each of which are the technical replicates). Far right column is the condition of the growth, either - or + ligand. For this particular switch, -L=repression and +L=expression. For those colonies that have less raw data values than the controls, the empty cells are filled with an integer that can be removed easily (-2) to be turned into NA when read into R. 
This is for repression and expression, fold induction will not be included because it will be calculated based on the included repression and expression values.
```{r}
#read in csv, these values will include colonies below the function cut off which has been defined here as 2.0
NH5_P3all <- read.csv("~/Downloads/Expression_platform_project/expression_platform_files/P3_library_data.csv", header=T)
empty_columns <- sapply(NH5_P3all, function(x) all(is.na(x) | x==""))
NH5_P3all <- NH5_P3all[,!empty_columns]
NH5_P3all[NH5_P3all == -2] <-NA
#MINUS 2AP DF
NH5_P3_minus1 <- NH5_P3all[NH5_P3all == "(-)",]
NH5_P3_minus <- NH5_P3_minus1[rowSums(is.na(NH5_P3_minus1)) !=ncol(NH5_P3_minus1),]
rm(NH5_P3_minus1)
#remove rows that were only NA and delete df that was filled with NA only columns
#PLUS 2AP DF
NH5_P3_plus1 <- NH5_P3all[NH5_P3all == "(+)",]
NH5_P3_plus <- NH5_P3_plus1[rowSums(is.na(NH5_P3_plus1)) !=ncol(NH5_P3_plus1),]
rm(NH5_P3_plus1)
#remove rows that were only NA 
```

###Using function to remove outliers and create a new dataframe in wide form and long form for minus
```{r}
noOut_NH5_P3_minus.list <- apply(NH5_P3_minus[,2:ncol(NH5_P3_minus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_NH5_P3_minus_sideways <- plyr::ldply(noOut_NH5_P3_minus.list, rbind)
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_NH5_P3_minus <- data.frame(t(noOut_NH5_P3_minus_sideways[-1]))
colnames(noOut_NH5_P3_minus) <- noOut_NH5_P3_minus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
NH5_P3_minus_noOut_wideform <- cbind(NH5_P3_minus[1:nrow(noOut_NH5_P3_minus),1], noOut_NH5_P3_minus[,1:ncol(noOut_NH5_P3_minus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(NH5_P3_minus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_NH5_P3_minus <- gather(NH5_P3_minus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
```

###Using function to remove outliers and create a new dataframe in wide form and long form for plus
```{r}

noOut_NH5_P3_plus.list <- apply(NH5_P3_plus[,2:ncol(NH5_P3_plus)], 2, removeOut, na.rm=TRUE)
#this used the function made to identify and remove outliers 
noOut_NH5_P3_plus_sideways <- plyr::ldply(noOut_NH5_P3_plus.list, rbind)
#THERE ARE TWO EMPTY ROWS HERE X AND X.1 THAT I WANT TO GET RID OF 
#took the list created from the applying the removeOut function and made it into a dataframe 
noOut_NH5_P3_plus <- data.frame(t(noOut_NH5_P3_plus_sideways[-1]))
colnames(noOut_NH5_P3_plus) <- noOut_NH5_P3_plus_sideways[,1]
#rotate the dataframe created by the function to prepare it for converting to long form 
NH5_P3_plus_noOut_wideform <- cbind(NH5_P3_plus[1:nrow(noOut_NH5_P3_plus),1], noOut_NH5_P3_plus[,1:ncol(noOut_NH5_P3_plus)])
#replaces id column with COLONY column which identified the value as (-) or (+)
colnames(NH5_P3_plus_noOut_wideform)[1] <- "COLONIES"
#rename the first column 
long_NH5_P3_plus <- gather(NH5_P3_plus_noOut_wideform, Colony_Name, Fluorescence, -COLONIES)
```

##Make DF of medians of each after removal of outliers 
```{r}
#median DFs
#medianP3_FI<- as.data.frame(apply(noOut_NH5_P3_FI, 2, median, na.rm=TRUE))
medianP3_minus<- as.data.frame(apply(noOut_NH5_P3_minus, 2, median, na.rm=TRUE))
medianP3_plus<- as.data.frame(apply(noOut_NH5_P3_plus, 2, median, na.rm=TRUE))
#standard error DFs for minus and plus
SE_P3_minus <- std.error(noOut_NH5_P3_minus)
SE_P3_plus <- std.error(noOut_NH5_P3_plus)
#create one data frame that has medians and standard errors
P3_all_summary_table <- cbind.data.frame(medianP3_minus, SE_P3_minus, medianP3_plus, SE_P3_plus)

#rename column names
colnames(P3_all_summary_table) <- c("-2AP", "-2AP_SE","+2AP", "+2AP_SE")
#make FI
P3_all_summary_table$FI <- P3_all_summary_table$`+2AP` / P3_all_summary_table$`-2AP`
#divide values by NH5
P3_all_summary_table$corr_minus_2AP <- as.numeric(P3_all_summary_table$'-2AP')/P3_all_summary_table[c("NH5"),c("-2AP")]
P3_all_summary_table$corr_plus_2AP <- as.numeric(P3_all_summary_table$'+2AP')/P3_all_summary_table[c("NH5"),c("+2AP")]
P3_all_summary_table$corr_FI <- as.numeric(P3_all_summary_table$FI)/P3_all_summary_table[c("NH5"),c("FI")]

#the measurements should have 2 sig figs and the standards errors should have 1 sig fig
two_sf_cols <- c(1,3,5,6,7,8)
one_sf_cols <- c(2,4)
P3_all_summary_table[,two_sf_cols] <- signif(P3_all_summary_table[,two_sf_cols],2)
P3_all_summary_table[,one_sf_cols] <- signif(P3_all_summary_table[,one_sf_cols],1)

#write csv
write.csv(P3_all_summary_table,"~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/P3/P3_all_summary_table.csv", row.names = TRUE)

#remove the switches below the functional cutoff of 2.0
#find the FI median
P3_summary_table <- subset(P3_all_summary_table, FI>1.9)
#write csv
write.csv(P3_summary_table,"~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/P3/P3_summary_table.csv", row.names = TRUE)
#write excel
write.xlsx(P3_summary_table, '~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/P3/P3_summary_table.xlsx', rowNames = TRUE)
write.xlsx(P3_all_summary_table, '~/Downloads/Expression_platform_project/expression_platform_files/analysis/output_files/P3/P3_all_summary_table.xlsx', rowNames = TRUE)

ggplot(P3_all_summary_table, aes(x=`corr_minus_2AP`, y=`corr_FI`))+ geom_point(size=2, alpha=1, show.legend = FALSE)+
  labs(title="P3 Library Repression vs Activity", x="Repression", y="Fold Induction")+
  theme_minimal()
```